---
title: "Reproducible Research: Peer Assessment 1"
author: "Hank C"
date: "9/20/2020"
output: 
  html_document:
    keep_md: true
---


```{r echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr,     warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)
library(zoo,       warn.conflicts = FALSE)

#read in data
activity <- read.table(file="activity/activity.csv", header=TRUE, sep=",")

# create factor for time intervals
activity$intervalFactor <- factor(activity$interval)

#create factor for date
activity$dateFactor <- factor(activity$date)

stepsByDate <- aggregate(activity$steps, by=list(Date=activity$dateFactor), FUN=sum, na.rm=TRUE)
names(stepsByDate) <- c("Date", "Steps")

stepsByInterval <- aggregate(activity$steps, by=list(Interval=activity$intervalFactor), FUN=sum, na.rm=TRUE)
names(stepsByInterval)<- c("Interval", "Steps")
```

### **Histogram of the total number of steps taken each day**
```{r echo=FALSE, warning=FALSE, message=FALSE}
dates           <- ymd(stepsByDate$Date)
dateRangeString <- paste( min(dates), " to ", max(dates), sep=" ")

plot(stepsByDate$Steps, type="s", main="Steps Per Day", xlab=dateRangeString, ylab="Steps", ylim = c(min(stepsByDate$Steps), max(stepsByDate$Steps)), xlim = c(1,length(stepsByDate$Steps) + 1))
```
<br><br>   

### **Time series plot of the average number of steps taken**
```{r echo=FALSE, warning=FALSE, message=FALSE}
weeklyMovingAverage <- transform(stepsByDate, roll = rollapplyr(Steps, 7, mean, partial = TRUE))
plot(weeklyMovingAverage$roll, type="l",  main = "7 Day Moving Average", xlab=dateRangeString, ylab="Steps")
```

### **Average Daily Activity**
#### There is more activity on weekends.

```{r echo=FALSE, warning=FALSE, message=FALSE}
daysOfWeek        <- wday(dates, label=TRUE)
stepsByDow        <- aggregate(stepsByDate$Steps, by=list(DoW=daysOfWeek), FUN=sum, na.rm=TRUE)
names(stepsByDow) <- c("DoW", "Steps")
totalDaysOfWeek   <- table(daysOfWeek)
avgStepsPerDoW    <- stepsByDow$Steps / totalDaysOfWeek

barplot(avgStepsPerDoW, type="s", ylim=c(0,1.25*max(avgStepsPerDoW)), ylab="Steps", main="Average Daily Activity", border="dark green", col="green")
```
```{r echo=FALSE, warning=FALSE, message=FALSE}
avgStepsPerDowSorted <- sort(avgStepsPerDoW, decreasing = TRUE)
barplot(avgStepsPerDowSorted, type="s", ylim=c(0,1.25*max(avgStepsPerDowSorted)), ylab="Steps", main="Average Daily Activity High to Low", border="dark green", col="green")
```

### **Mean and median number of steps taken each day**
```{r echo=FALSE, warning=FALSE, message=FALSE}

stepsByDate_mean   <- mean(stepsByDate$Steps)
stepsByDate_median <- median(stepsByDate$Steps)
```
#### Number of steps taken per day (mean)  : `r format(round(stepsByDate_mean, 1), nsmall = 1)`
#### Number of steps taken per day (median): `r stepsByDate_median`
<br>
     

### **Most Active 5 Minute Interval**
```{r echo=FALSE, warning=FALSE, message=FALSE}
totalIntervals        <- table(activity$interval)
avgStepsPerInterval   <- stepsByInterval$Steps / totalIntervals
intervalMax           <- max(avgStepsPerInterval)
intervalMaxIndex      <- avgStepsPerInterval[which.max(avgStepsPerInterval)]
intervalMaxName       <- names(intervalMaxIndex)
```
####  
#### Average steps of the most active interval: `r format(round(intervalMax, 1), nsmall = 1)`
#### Most active interval: `r intervalMaxName`


### **Code to describe and show a strategy for imputing missing data**
```{r echo=FALSE, warning=FALSE, message=FALSE}

# see if there are any missing dates
dates         <- as.Date(levels(activity$dateFactor))
dates_range   <- seq(min(dates), max(dates), by=1)
missing_dates <- dates_range[!dates_range %in% dates]

activitySummary <- summary(activity$steps)
```
#### Number of missing dates : `r length(missing_dates)`
#### Number of NA step values: `r activitySummary["NA's"]`
<br>   

#### The data contains no missing dates.
#### Imputation will be restricted to replacing NA values with 0.
<br><br>     

### **Histogram of the total number of steps taken each day after missing values are imputed**
```{r echo=FALSE, warning=FALSE, message=FALSE}
#replace NA steps with 0
activity$steps[is.na(activity$steps)] <- 0

stepsByDate <- aggregate(activity$steps, by=list(Date=activity$dateFactor), FUN=sum, na.rm=TRUE)
names(stepsByDate) <- c("Date", "Steps")

plot(stepsByDate$Steps, type="s", main="Steps Per Day (Imputed)", xlab=dateRangeString, ylab="Steps", ylim = c(min(stepsByDate$Steps), max(stepsByDate$Steps)), xlim = c(1,length(stepsByDate$Steps) + 1))

```


